---
title: "R Notebook"
output: html_notebook
---



```{r}
library(fpp)
library(fpp2)
```

*** Retail debit card usage in iceland ***

```{r}
tranformdebit=debitcards

```

*** Time Series Data ***


```{r}
class(tranformdebit)
```
Exploring the data
*** plotting the data ***

```{r}
autoplot(transformdebit) + ggtitle("Retail debit card usage in Iceland") + xlab("Year") + ylab("million ISK")

```




```{r}
ggsubseriesplot(transformdebit) + ylab("million ISK") + ggtitle("Seasonal subseries plot:Retail debit card usage in Iceland ")
```

Autocorrelation
```{r}
ggAcf(transformdebit)
```



*** Decomposing the data to understand trend and seasonality ***

```{r}
decompose<-stl(transformdebit,s.window = 5)#(check)
plot(decompose)
```
```{r}
autoplot(transformdebit)+
  autolayer(decompose$time.series[,2],series="Trend")+
   ggtitle("Retail debit card usage in Iceland") + 
  xlab("Year") + ylab("million ISK")+
  guides(colours=guide_legend(title='Trend'))
```

```{r}
#seasonal component over time
monthplot(decompose$time.series[,"seasonal"], main="", ylab="Seasonal")
```






*** Splitting the data into training and test ***

```{r}

train<- window(transformdebit,start=2000,end=c(2008,12),frequency=12)
test<-window(transformdebit,start=c(2009,1),end=c(2012,12),frequency=12)

```

*** Forecasting Retail debit card usage using Average method ***

```{r}
fit_mean<-meanf(train,h=48)
autoplot(window(train,start=2000,end=))+autolayer(fit_mean,series="mean",PI=FALSE)+xlab("Years")+ylab("Usage")+ggtitle("Retail debit card usage in Iceland")+guides(colour=guide_legend(title = "Forecast"))
accuracy(fit_mean,test)

```

*** Forecasting Retail debit card usage using Naive method (since it is widely used for economic and finanicial time series)***

```{r}
fit_naive<-naive(train,h=48)
autoplot(window(train,start=2000,end=))+autolayer(fit_naive,series="naive",PI=FALSE)+xlab("Years")+ylab("Usage")+ggtitle("Retail debit card usage in Iceland")+guides(colour=guide_legend(title = "Forecast"))
accuracy(fit_naive,test)
```
*** Forecasting Retail debit card usage using Seasonal Naive method, since there is seasonality in the data ***

```{r}
fit_snaive<-snaive(train,h=48)
autoplot(window(train,start=2000))+autolayer(fit_snaive,series="snaive",PI=FALSE)+xlab("Years")+ylab("Usage")+ggtitle("Retail debit card usage in Iceland")+guides(colour=guide_legend(title = "Forecast"))
accuracy(fit_snaive,test)
```
*** Forecasting Retail debit card usage using drift method ***

```{r}
fit_drift<-rwf(train,h=12,drift=TRUE)
autoplot(window(train,start=2000))+autolayer(fit_drift,series="Drift",PI=FALSE)+xlab("Years")+ylab(" Amount accumalted by Debit Card in millions ISK")+ggtitle("Retail debit card usage in Iceland")+guides(colour=guide_legend(title = "Forecast"))
accuracy(fit_drift,test)
```
*** Forecasting Retail debit card usage using Holt winter's additive method ***

```{r}
fit_hwa<-hw(train,h=48,seasonal='additive')
autoplot(window(train,start=2000,end=))+autolayer(fit_hwa,series="HW Additive Forecasting",PI=FALSE)+xlab("Years")+ylab("Usage")+ggtitle("Retail debit card usage in Iceland")+guides(colour=guide_legend(title = "Forecast"))
accuracy(fit_hwa,test)

```
*** Forecasting Retail debit card usage using Holt winter's multiplicative method ***

```{r}
fit_hwa<-hw(train,h=48,seasonal='multiplicative')
autoplot(window(train,start=2000,end=))+autolayer(fit_hwa,series="HW Multiplicative Forecasting",PI=FALSE)+xlab("Years")+ylab("Usage")+ggtitle("Retail debit card usage in Iceland")+guides(colour=guide_legend(title = "Forecast"))
accuracy(fit_hwa,test)

```

*** Forecasting Retail debit card usage using Arima method ***

```{r}
fit_arima<-auto.arima(train)
forecast_arima<-forecast(fit_arima,h=12)
autoplot(window(train,start=2000,end=))+autolayer(forecast_arima,series="Arima Forecasting",PI=FALSE)+xlab("Years")+ylab("Usage")+ggtitle("Retail debit card usage in Iceland")+guides(colour=guide_legend(title = "Forecast"))
accuracy(forecast_arima,test)

```



